---
title: "JMU_Week4"
author: "Joanne Unite"
output: word_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
#Downloading packages
```{r}
library(readr)
library(tidyr)
library(dplyr)
library(ggplot2)
```

#This first dataset is from the CDC website and documents the COVID-19 cases or deaths by age group and vaccination status

```{r}
cdc_df <- read.csv("/Users/junite/Desktop/BIFX551/BIFX551_SPRING24/Week 4/covid19.csv")
```

##Look at the specs of the data
```{r}
summary(cdc_df)
class(cdc_df)
```
>It looks like there are many NA values throughout the data frame so if I choose to use this data to use for my final project I would need to clean up those values. 

###Try an experimental plot of the data
```{r}
ggplot(cdc_df, aes( x = Vaccine.product)) +
  geom_bar()
```
> This plot shows us that the vast majority of cases documented included groups that had a mix of the vaccine brands. I would like to explore the amount of cases in these four groups and how it breaks down by age group.

***

#This next data set is from Kaggle which contains clinical data from patients to assess cardiovascular health and kidney function

```{r}
blood_df <- read.csv("/Users/junite/Desktop/BIFX551/BIFX551_SPRING24/Week 4/blood_test.csv")
```

## Let's take a look at the data
```{r}
summary(blood_df)
class(blood_df)
```

### Try an experimental plot on the data

```{r}
ggplot(blood_df,aes(x = Age, y = Chol)) +
  geom_point()
```
> This preliminary plot shows us that many of the patients fall within range of 3 to 6.5 for their cholesterol levels from blood samples. We can also see several outliers at varying ages. If I were to choose this dataset for my project I would like to explore the break down of each test by males and females. I would explore what the most important tests in the dataset that would give the greatest indication of risk of heart disease and diabetes.

***

#This last dataset is from data.world which documents health data from major cities across the USA.

```{r}
cities_df <- read.csv("/Users/junite/Desktop/BIFX551/BIFX551_SPRING24/Week 4/Big_Cities.csv")
```

## Let's take a look at the summary of the data
```{r}
summary(cities_df)
class(cities_df)
```

### Try a preliminary plot with the data
```{r}
ggplot(cities_df, aes(x = Gender)) + 
  geom_bar()
```

> This plot shows me that I will not be able to break down/explore the data by gender since the majority of the data is categroized as "both". Based off the prelimenary look at the data, this data will need a lot of cleaning. The data contains columns from the same city with the same health indicator category but for some of the columns there is an "age-adjusted" indector. I am not sure what that means and would likely need to creat subsets of the data to analyze.