---
title: "Spatial Data/Maps"
author: Dina Lauricella
date: "`r format(Sys.Date())`"
output:
  html_document:
    toc: yes
    toc_float: yes
    codefolding: show
    number_sections: no
  word_document:
    toc: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(readr)
library(dplyr, quietly = TRUE)
library(ggplot2)
library(RColorBrewer)
library(sf)
library(maps)
```
## Data and data structure:

```{r}
# https://simplemaps.com/data/us-zips
zip_csv <- read_csv("/Users/dml/Documents/GitHub/BIFX551_SPRING24/Week 9/uszips.csv")

# View data
str(zip_csv)
```

## US outline map
```{r}
us_map <- map_data("usa" )
ggplot( aes(x = long, y = lat), data = us_map ) +
  geom_polygon( aes(group = group), fill=NA, color="blue") +
  coord_quickmap()
```

## US zip code density map
From this you can see that the zip codes correspond to the population density areas around big cities. You can also see horizontal lines of zip codes in the Midwest. If there were major highway data available, it would be interesting to see if the zip codes follow the major roads.
```{r}
# Load map data for the USA
usa_map <- map_data("usa")

# Load state boundaries
state_lines <- map_data("state")

# Plot zip code on map

# Plot the map with latitude and longitude data
ggplot() +
  geom_polygon(data = usa_map, aes(x = long, y = lat, group = group), fill = "grey90") +
  geom_path(data = state_lines, aes(x = long, y = lat, group = group), color = "blue") +
  geom_point(data = zip_csv, aes(x = lng, y = lat), color = "red", alpha = .1, shape = 20) +
  coord_cartesian(xlim = c(-130, -70), ylim = c(25, 50)) +
  labs(x = "Longitude", y = "Latitude", title = "US Zip Codes")

```

## Michigan outline and zip code densities map
It is possible that the zip codes in Michigan follow major roads as well.
```{r}
mi_map <- map_data("state", "Michigan" )

# Read us zip codes file
zip_csv <- read_csv("/Users/dml/Documents/GitHub/BIFX551_SPRING24/Week 9/uszips.csv")

# Filter zip code data for Michigan only
zip_csv_mi <- zip_csv[zip_csv$state_name == "Michigan", ]

ggplot( aes(x = long, y = lat), data = mi_map ) +
  geom_polygon( aes(group = group), fill=NA, color="lightblue") +
  coord_quickmap()

# Plot the map of Michigan
d <- ggplot() +
  geom_polygon(data = mi_map, aes(x = long, y = lat, group = group), fill = "grey95", color = "lightblue") +
  coord_fixed() +  # Keep aspect ratio
  labs(x = "Longitude", y = "Latitude", title = "Michigan Map with Zip Codes")

# Overlay zip code points for Michigan
if (nrow(zip_csv_mi) > 0) {
  d <- d +
    geom_point(data = zip_csv_mi, aes(x = lng, y = lat), color = "blue", alpha = .1) +
    labs(x = "Longitude", y = "Latitude", title = "Michigan Zip Codes")
  } else {
  print("No zip code data for Michigan provided.")
}

print(d)

```

## Maryland outline and zip code densities map
It is harder to see if zip codes in Maryland follow major roads. Unlike the Midwest, Eastern roads are curvier, so straight lines are not seen.
```{r}
md_map <- map_data("state", "Maryland" )

# Read us zip codes file
zip_csv <- read_csv("/Users/dml/Documents/GitHub/BIFX551_SPRING24/Week 9/uszips.csv")

# Filter zip code data for Maryland only
zip_csv_md <- zip_csv[zip_csv$state_name == "Maryland", ]

ggplot( aes(x = long, y = lat), data = md_map ) +
  geom_polygon( aes(group = group), fill=NA, color="purple") +
  coord_quickmap()

# Plot the map of Maryland
p <- ggplot() +
  geom_polygon(data = md_map, aes(x = long, y = lat, group = group), fill = "grey95", color = "purple") +
  coord_fixed() +  # Keep aspect ratio
  labs(x = "Longitude", y = "Latitude", title = "Maryland Map with Zip Codes")

# Overlay zip code points for Maryland
if (nrow(zip_csv_md) > 0) {
  p <- p +
    geom_point(data = zip_csv_md, aes(x = lng, y = lat), color = "black", alpha = .1) +
    coord_cartesian(xlim = c(-79.8, -75), ylim = c(37.8, 39.8)) +
    labs(x = "Longitude", y = "Latitude", title = "Maryland Zip Codes")
  } else {
  print("No zip code data for Maryland provided.")
}

print(p)
```



