---
title: "MJM_Week9"
author: "Mike Murphy"
date: "r Sys.Date()"
output: html_document
---
# Loading librarys
```{r}
library(ggplot2)
library(maps)
library(leaflet)
library(readxl)
library(sf)
```

# Download and load CSV file
```{r}
# CSV from https://simplemaps.com/data/us-zips

zipcodes <- read.csv("uszips.csv")
```

# Load map data and generate a blank map of the US
```{r}
usa_map <- map_data("state") #get a map of the US states  
USPlot <- ggplot( aes(x=long, y=lat ), data = usa_map ) +  
   geom_polygon( aes(group = group), fill = NA, show.legend = FALSE, color="grey50") +  
  coord_quickmap() + 
  theme_void()

USPlot
```

```{r}
uszipcodes <- zipcodes %>% 
  filter(between(lat, 25, 50) & between(lng, -127, -65 ))

zipcode_plot <- USPlot + geom_point(aes(x = lng, y= lat), data = uszipcodes, alpha = 0.05)

zipcode_plot
```
# What can you infer from this plot?
## We can infer that the darker zones caused from more points overlapping consist of more zipcodes, and the lighter areas consist of less overlapping points and fewer zipcodes. 
# Does the concentration of zip codes in a given area tell you anything about that area?
## Concentration of zipcodes is a good measure of population density.
# What would be the next steps to test your hypothesis?
## We can plot the US cities on the map and see if they coincide with the darker, or more overlapped points of zipcodes. 
```{r}
us_cities <- us.cities

#Sort the cities by population
us_cities <- us_cities[order(us_cities$pop),]

#Take the top 25 cities by population
topcities <- tail(us_cities, n=25)

#Plot the top 25 cities
zipcode_plot + 
  geom_point(aes(x = long, y= lat, color = name), data = topcities, size = 4, alpha = 0.5)
```
# We can see that the cities sit on top of a lot of the areas with more zipcodes. 
