---
title: "RPackage"
author: "Jamilet Cordon"
date: "2024-04-01"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{RPackage}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>")
```

## Writing an R package from Scratch

```{r packages}
#install.packages("devtools")
#install.packages("roxygen2")
library(roxygen2)
library(devtools)

#Create package directory 
source("../R/Celsius_Conversions.R")
source("../R/Fahrenheit_Conversions.R")
```

### Question 1
## Function

##  Fahrenheit to Celsius 
```{r, error=TRUE}

F_to_C(56) 
```
## Celsius to Fahrenheit 
```{r, error=TRUE}

C_to_F(23) 
```
## Question 2

Under the Test folder I can see that the files are labeled with its intended function.

The gather package has multiple functions that tells the program to search/collect for data

within columns of tables. There are functions for errors, groups, empty columns,

integers vs character etc. In this package there are some functions I do not understand, but

I think its due to not knowing what is in the datasets. There is a test that, "gathers throws

error for werid objects", it is hard to compute b/c I do not know what or how the program

will know what "werid objects" are. It appears to mean that the packages under test are

like The packages under Vignett are written in a style that is familar .Rmd. These are the

code/function that we have mainly performing. With a written preview and a code preview.

there is no data available

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
