---
title: "DML_week5"
author: "Dina Lauricella"
date: "`r format(Sys.Date())`"
output:
  word_document:
    toc: yes
  html_document:
    toc: yes
    toc_float: yes
    codefolding: show
    number_sections: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Import Libraries

```{r}
library(readr)
library(tidyr)
library(dplyr)
library(janitor)
library(ggplot2)
```

## Read Opiod Data (1999-2014)

```{r}
opiod <- read_csv("C:/Users/dlauricella/Documents/GitHub/BIFX551_SPRING24/Week 4/Opiod Death, 1999-2014 v1.1.csv", "opiod.csv")

```



## Take a look at Opiod data

```{r}
opiod2 <- janitor::row_to_names(opiod, 1)
opiod2 <- opiod2 %>% rename("num_Rx_dispensed" = "Prescriptions Dispensed by US Retailers in that year (millions)")
class(opiod2)
str(opiod2)
spec(opiod2)
spec(opiod)
summary(opiod2)
dim(opiod2)
sum(is.na(opiod2))
```

```{r}
west <- c("Washington", "Oregon", "California", "Alaska", "Montana", "Idaho", "Wyoming", "Nevada", "Utah", "Colorado", "Arizona", "New Mexico", "Hawaii")
midwest <- c("North Dakota", "South Dakota", "Nebraska", "Kansas", "Minnesota", "Iowa", "Missouri", "Wisconsin", "Illinois", "Michigan", "Indiana", "Ohio")
south <- c("Texas", "Oklahoma", "Arkansas", "Louisiana", "Kentucky", "Tennesse", "Mississippi", "Alabama", "West Virginia", "Virginia", "North Carolina", "South Carolina", "Georgia", "Florida")
northeast <- c("New York", "Pennsylvania", "Maine", "Vermont", "New Hampshire", "Massachusetts", "Rhode Island", "Connecticut", "New Jersey", "Delaware", "Maryland")

```


## Plot Opiod Data with Contious Variables
## This shows that deaths increased with increasing prescriptions dispensed.

```{r}

# Remove rows with NA values in the 'Deaths' column
opiod2 <- opiod2 %>% filter(!is.na(Deaths))

# Convert 'Deaths' and 'num_Rx_dispensed' columns to numeric
opiod2$Deaths <- as.numeric(opiod2$Deaths)
opiod2$num_Rx_dispensed <- as.numeric(opiod2$num_Rx_dispensed)

# Group by Year and calculate the sum of Deaths and Prescriptions
opiod_sum <- opiod2 %>%
  group_by(Year) %>%
  summarise(TotalDeaths = sum(Deaths, na.rm = TRUE),
            Total_num_Rx_dispensed = sum(num_Rx_dispensed, na.rm = TRUE))

# Plot
ggplot(opiod_sum, aes(x = Total_num_Rx_dispensed, y = TotalDeaths)) +
  geom_point() +
  labs(x = "Total Prescriptions Dispensed (Millions)",
       y = "Total Deaths",
       title = "Deaths as a Function of Prescriptions by Year (1999-2014)")

```


## Plot Opiod Data with Discrete Variables
## The greater the population size, the greater the deaths.

```{r}

# Convert 'Deaths' and 'Population' columns to numeric
opiod2$Deaths <- as.numeric(opiod2$Deaths)
opiod2$Population <- as.numeric(opiod2$Population)

# Group by State and calculate the mean of Population and Deaths
opiod_mean <- opiod2 %>%
  group_by(State) %>%
  summarise(MeanPopulation = mean(Population, na.rm = TRUE),
            MeanDeaths = mean(Deaths, na.rm = TRUE))

# Plot
ggplot(opiod_mean, aes(x = MeanPopulation, y = MeanDeaths)) +
  geom_point() +
  labs(x = "Mean Population by State",
       y = "Mean Deaths by State",
       title = "Mean Deaths as a Function of Mean Population by State")


```


## Plot Opiod Data with Color/Size to Represent Quantity
## This shows death by region normalized to the population size. By region is not as dramatic as individual states would have shown.

```{r}

# Convert 'Deaths' column to numeric
opiod2$Deaths <- as.numeric(opiod2$Deaths)

# Assign states to regions
opiod2 <- mutate(opiod2,
                 Region = case_when(
                   State %in% c("Washington", "Oregon", "California", "Alaska", "Montana", "Idaho", "Wyoming", "Nevada", "Utah", "Colorado", "Arizona", "New Mexico", "Hawaii") ~ "West",
                   State %in% c("North Dakota", "South Dakota", "Nebraska", "Kansas", "Minnesota", "Iowa", "Missouri", "Wisconsin", "Illinois", "Michigan", "Indiana", "Ohio") ~ "Midwest",
                   State %in% c("Texas", "Oklahoma", "Arkansas", "Louisiana", "Kentucky", "Tennessee", "Mississippi", "Alabama", "West Virginia", "Virginia", "North Carolina", "South Carolina", "Georgia", "Florida") ~ "South",
                   State %in% c("New York", "Pennsylvania", "Maine", "Vermont", "New Hampshire", "Massachusetts", "Rhode Island", "Connecticut", "New Jersey", "Delaware", "Maryland") ~ "Northeast",
                 ))

# Filter out rows with NA in the Region column
opiod2 <- opiod2 %>% filter(!is.na(Region))

# Group by region and calculate total deaths and total population for each region
opiod_region <- opiod2 %>%
  group_by(Region) %>%
  summarise(TotalDeaths = sum(Deaths, na.rm = TRUE),
            TotalPopulation = sum(Population, na.rm = TRUE))

# Calculate deaths per 100,000 population
opiod_region <- mutate(opiod_region, DeathsPer100k = (TotalDeaths / TotalPopulation) * 100000)

# Create a bar plot with deaths per 100,000 population by region
ggplot(opiod_region, aes(x = Region, y = DeathsPer100k, fill = DeathsPer100k)) +
  geom_bar(stat = "identity") +
  scale_fill_viridis_c() +  # You can change the color scale as needed
  labs(x = "Region",
       y = "Deaths per 100,000 Population",
       title = "Opioid Deaths per 100,000 Population by Region") +
  theme_minimal()


```


## Plot Opiod Data as Faceted Plot
## The plot showing the highest deaths/year shows a steeper rise than the lowest deaths/year.

```{r}

# Convert 'Deaths' and 'num_Rx_dispensed' columns to numeric
opiod2$Deaths <- as.numeric(opiod2$Deaths)
opiod2$num_Rx_dispensed <- as.numeric(opiod2$num_Rx_dispensed)

# Identify the years with the lowest and highest total deaths
lowest_death_year <- opiod2 %>%
  group_by(Year) %>%
  summarise(TotalDeaths = sum(Deaths, na.rm = TRUE)) %>%
  slice_min(order_by = TotalDeaths) %>%
  pull(Year)

highest_death_year <- opiod2 %>%
  group_by(Year) %>%
  summarise(TotalDeaths = sum(Deaths, na.rm = TRUE)) %>%
  slice_max(order_by = TotalDeaths) %>%
  pull(Year)

# Filter data for the selected years
opiod2_filtered <- opiod2 %>%
  filter(Year %in% c(lowest_death_year, highest_death_year))

# Create a faceted plot for the selected years
ggplot(opiod2_filtered, aes(x = Population, y = Deaths, color = as.factor(Year))) +
  geom_point(alpha = 0.7) +
  labs(x = "Population by State",
       y = "Total Deaths by State",
       title = "Deaths, Population, and Prescriptions by State (Lowest and Highest Death Years)") +
  facet_wrap(~ Year, scales = "free_y", ncol = 2) +
  scale_color_viridis_d() +
  theme_minimal()


```

