---
title: "Week 9"
author: "Josie Sdhultz"
date: "2024-03-28"
output: html_document
---

Functioning map for reference
Use maps package, seperate by state and create gray borders

```{r}
library(ggplot2)

install.packages("maps")
library(maps)  
usa_map <- map_data("state") #get a map of the US states  
ggplot( aes(x=long, y=lat ), data = usa_map ) +  
  geom_polygon( aes(group = group, fill=region), show.legend = FALSE, color="grey50") +  
  coord_quickmap()
```

Had to download file containing zip code data manually and use it since Kaggle is problematic. 

URL for US Zip code data; https://www.kaggle.com/datasets/jedwible/uszipcodes-20231227

File name; USZipsWithLatLon_20231227.csv

Skipped country code column since it is all US. 
```{r}
USZipsWithLatLon_20231227.csv <- "D:/Downloads/zip/USZipsWithLatLon_20231227.csv"

library(readr)
zipcode<- read_csv((USZipsWithLatLon_20231227.csv), skip=1, col_names = c("country_code", "postal_code", "place_name", "admin_name1", "admin_code1", "admin_name2", "admin_code2", "latitude", "longitude"),
col_types=cols(country_code=col_skip(), postal_code=col_integer(), place_name=col_character(), admin_name1=col_character(), admin_code1=col_integer(), admin_name2=col_character(), admin_code2=col_integer(), latitude=col_double(), longitude=col_double()))

```




Made the same data from us map, but specified for Pennsylvania. Filtered zipcode data to search for only Pennsylvania. Plotted pazip and plotpamap and adjusted the intensity of alpha to see dense populations easily. 
```{r}
library(ggplot2)
library(dplyr)
library(maps)  

pamap <- map_data("state", "Pennsylvania") 

plotpamap <- ggplot( aes(x=long, y=lat ), data = pamap ) +  
  geom_polygon( aes(group = group), fill="magenta", show.legend = FALSE, color="grey50") +  
  coord_quickmap()

pazip <- filter(zipcode, admin_name1== "Pennsylvania")

plotpamap + geom_point(aes(x=longitude, y=latitude), shape = 20, data = pazip, alpha=.3)
```
1. The more zipcodes, the greater the population. Therefore the amount of zipcodes is indicative of population density.  
2. The concentration of zip codes indicates how highly populated it is. The more overlap/darker the plotting, the more dense in population the area is.
3. Make a heatmap to show variation of zipcodes across the state. Having another heatmap based off of population itself would be beneficial to compare one another. 
