---
---
---

## AJO_Week5 Homework Assignment

------------------------------------------------------------------------

### Data Citation : Heart Disease Dataset Available on Kaggle.com

## URL: <https://www.kaggle.com/datasets/yaminh/heart-disease-dataset/data>

```{r}

library(dplyr)
library(tidyr)
library(ggplot2)

```

### Loading packages and reading data frame + summary

[STEPS]{.underline}

1.) Find the dataset you want to use for midterm and this assignment.

2.) Make folder with the assignment in it and save the .csv to that folder.

3.) Read the dataset into R studio and begin working on assignment

```{r}
# Bring in the heart data

heart_df <- read.csv("heart_data.csv")

```

#### a.) A plot with continuous variables

```{r}
# Continuous Variables - age and maximum heart rate
# Make df for age and max_hr
age_max_hr_df <- select(heart_df, age, max_hr)

# Create the plot with trend line
ggplot(age_max_hr_df, aes(x = age, y = max_hr)) +
  geom_point() +
  geom_smooth(method = "lm", formula = 'y ~ x', se = FALSE) +  
  labs(x = "Age", y = "Max Heart Rate", caption = "Figure 1: This scatter plot shows the patient's age and maxiumum heart rate recorded.\n The trend line shows the relationship between age and maximum heart rate.") +
  ggtitle("Max Heart Rate of Patients by Age") +
  theme(plot.caption = element_text(hjust = 0.75)) +
  theme(plot.title = element_text(hjust = 0.5)) 
  
```

#### b.) A plot where you use color or size to represent a quantity

```{r}
# Discrete Variables - Diagnosis and Gender

#Factor variables 

# Convert diagnosis to factor variable
heart_df$diagnosis <- factor(heart_df$diagnosis, levels = c(0, 1), labels = c("No Heart Disease", "Diagnosed with Heart Disease"))

# Convert Gender to factor variable
heart_df$gender <- factor(heart_df$gender, levels = c(0, 1), labels = c("Male", "Female"))


ggplot(heart_df, aes(x = gender, y = diagnosis)) +
  geom_bin_2d() +
  labs(x = "Sex", y = "Diagnosis", size = "Number of Patients" ) +
  labs(caption = "Figure 2: This heatmap shows the number of patients (count)\n with or without a heart disease diagnosis by their sex.") +
  ggtitle("Diagnosis of Heart Disease by Patient Sex") +
  theme(plot.caption = element_text(hjust = 0.75)) +
  theme(plot.title = element_text(hjust = 0.5)) +
  scale_fill_gradient(low = "yellow", high = "red")


```

\

#### c.) A plot with discrete variables

```{r}

# Create age groups

# Define the breaks for age groups
breaks1 <- c(0, 50, max(heart_df$age))

# Define labels for the age groups
labels1 <- c("0-50","51+")

# Create the age_groups column
heart_df$age_group <- cut(heart_df$age, breaks = breaks1, labels = labels1, right = TRUE)
 
# factor age group

heart_df$age_group <- factor(heart_df$age_group)


# Calculate the average chest_pain score for each age_group
average_chest_pain <- heart_df %>%
  group_by(age_group) %>%
  summarize(average_chest_pain = mean(chest_pain))

  
# Create the col bar plot
ggplot(average_chest_pain, aes(x = age_group, y = average_chest_pain)) +
  geom_col(fill = "blue") +
  labs(x = "Age Group", y = "Average Chest Pain", caption = 
"Figure 3: This bar graph shows the average result from heart pain scale reports from adult patients within various age groups.\n The chest pain scale cosists of 0 (no pain) to 3 (intolerable). ") +
  ggtitle("Average Chest Pain Reported by Adult Patients") +
  theme(plot.caption = element_text(hjust = 0.5)) + 
  theme(plot.title = element_text(hjust = 0.5)) 

```

#### d.) A faceted plot

```{r}



# Create the age_groups column
heart_df$age_group <- cut(heart_df$age, breaks = breaks1, labels = labels1, right = TRUE)
 
# factor age group

heart_df$age_group <- factor(heart_df$age_group)


#make new column
heart_df <- heart_df %>% mutate(risk_score = age + chest_pain + cholesterol + st_depression)

# Calculate the average risk_score for each gender and age_group
average_risk_score <- heart_df %>%
  group_by(gender, age_group) %>%
  summarize(average_risk_score = mean(risk_score))

# Create the faceted bar plot
ggplot(average_risk_score, aes(x = gender, y = average_risk_score, fill = gender)) +
  geom_col() +
  facet_wrap(~ age_group, scales = "free") +
  labs(x = "Gender", y = "Average Risk Score", fill = "Gender", caption = "Figure 4: This faceted column bar plot compares the average calculated risk score\n between males and females within two age groups,\n 0-50 years of age, and 51+ years of age.") +
  ggtitle("Average Risk Score by Gender and Age Group") +
  theme(plot.title = element_text(hjust = 0.5)) +
  theme(plot.caption = element_text(hjust = 0.5)) 

```

### Data Citation : Heart Disease Dataset Available on Kaggle.com

## URL: <https://www.kaggle.com/datasets/yaminh/heart-disease-dataset/data>
