---
title: "JMU_Week9"
author: "Joanne Unite"
date: "2024-03-26"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

# install packages

library(dplyr, quietly = TRUE)
library(ggplot2)
library(RColorBrewer)
library(sf)
library(maps)
library(tidyr)
```

# Uploading Data

I retrieved the data from Kaggle: https://www.kaggle.com/datasets/pavansanagapati/us-wages-via-zipcode

```{r}
zips <- read.csv("/Users/junite/Desktop/BIFX551/BIFX551_SPRING24/Week 9/JMU_Week 9/zipcode-database.csv")
```


# Inspect the data
```{r}
str(zips) 
```
# Explore Data

```{r, explore}
#inspect the number of zipcodes per state
zips %>% 
  group_by( State ) %>% 
  summarise(nZipcode = n_distinct(Zipcode) )

#inspect the amount of cities per state
zips %>% 
  group_by( State ) %>%
  summarise(nCity = n_distinct( City) )

```





```{r, creating map with states }
#use map_data to select what we want
states_map <- map_data("state")

#look at the new states data frame
View(states_map)

#Plot the map to check if it looks good
states_map %>%
  ggplot( aes(x = long, y = lat) ) +
  geom_polygon( aes(group = group), fill= NA, color="black") +
  theme_void()+
  theme(legend.position = "none") +
  coord_quickmap()


```

```{r, Zipcode location plot}
#Use the map created above and add a layer with zipcode points
ggplot(zips, aes(Long,Lat), fill = State) +
  geom_polygon(data = states_map,aes(x = long,y = lat, group = group, fill = region),color='blue', show.legend = FALSE, alpha=.35) +
  geom_point(aes(),size=.15,alpha=.25) +
  xlim(-125,-65) + ylim(22,50) + 
  ggtitle("Zipcode Density Across the United States") + 
  coord_quickmap()
```

# Assignment questions

## What can you infer from this plot?

You can infer that the highly dense areas are likely where population is higher.

## Does the concentration of zip codes in a given area tell you anything about that area?

You can observe that the there is a higher number of points located surrounding major cities. 

## What would be the next steps to test your hypothesis?

The next step I would do to test my hypothesis is to create a map showing the major cities across the United states. Then I would create a density plot on the map using the population data.
